Towards Reproducibility in ANS Performance

### 2019-10-04 paper.tex
final code version

\hypersetup{
            pdftitle={Busting the Myth! - Reproducibility in ANS Performance},
            pdfauthor={Enrico Spinielli, Rainer Koelle},
            pdfkeywords={reproducibility, ANS performance, flight trajectory, open data, performance indicator},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Busting the Myth! - Reproducibility in ANS Performance}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Air transportation is undergoing a fundamental transformation. On the
political, strategic, and operational level work is directed to meet the
future growth of air traffic. To meet this goal, operational concepts
and technological enablers are deployed with the promises of distinct
performance improvements. These improvements are typically widely
marketed, however, the underlying evidence is not made available. From
that respect Air Navigation System Performance suffers from the same
credibility crisis seen in other industries. This paper implements the
reproducibility paradigm by providing data, code, and results openly to
enable another researcher or analyst to reproduce - and potentially -
improve or build on our work. This work is part of a wider thread to
establish Open Data and Open Source based data analytical capabilities
for Air Navigation System Performance in Europe. The implementation of
the paradigm is demonstrated on the basis of the performance reference
trajectory by replicating a classical performance measure, and then
moving into investigating the application of trajectory based
information for enhancing the current state of performance monitoring
for the arrival phase. This paper applies the concept and approach as an
use case analysis for 2 European airports with different strategies for
sequencing arrival traffic (i.e.~stack holding and point merge). The
work reported demonstrates the feasibility of the approach and reflects
a stepping stone in the roll-out of the performance reference trajectory
within the European context. The work also builds the basis for novel
developments on the basis of the implemented reproducibility workflow as
this paper is accompanied by a webpage that supports the impartial
evaluation of the work by providing the data, code, and how this reports
had been produced.
\end{abstract}

% keywords
\begin{IEEEkeywords}
\textit{%
reproducibility; ANS performance; flight trajectory; open data; performance indicator}
\end{IEEEkeywords}

@@

@@

\section{Introduction}\label{introduction}

Air transportation is undergoing a fundamental transformation. To meet
the future growth of air traffic, political decision makers, strategic
planers, practitioners and researchers aim to enhance the performance of
air navigation services (ANS). At the time being there is no consistent
and regionally harmonised deployment programme. Typically, the
deployment programmes are revolving around public-private partnerships
and regional funding programmes resulting in a wide mix of consortia and
stakeholder subsets. The actual deployment of novel concepts or
technological enablers is then driven by the programmes and to a lesser
extent by operational network wide priorities following a piece meal
approach.

Moreover, the success of these deployments is typically supported by
promising business cases and operational evaluations - often expressed
as operational performance improvements. While these improvements are
marketed, the underlying data and associated analyses in support of the
results are seldomly made available. Explanations range from data
sensitivity, commercial interests and property rights, through to
complexity of the data analyses or simulations, and the associated
volume of data.

Accordingly, ANS performance improvements are subject to a ``credibility
crisis'' similar to reported problems in other domains, e.g.~medicine
{[}\protect\hyperlink{ref-kolata_2011}{1}{]}, psychology
{[}\protect\hyperlink{ref-ritchie_2012}{2}{]}, ecology
{[}\protect\hyperlink{ref-fraser_etal_2018}{3}{]}. Fidler and Wilcox
(2018) provides a recent review of the reproducibility discussion
{[}\protect\hyperlink{ref-fidler_wilcox_2018}{4}{]}. Part of the crisis
sensation was driven by the diffusion of reported results as other
researchers/analysts were able to discover substantial flaws in the data
analyses. However, a crisis also offers room for self-correction.
Throughout the recent years, this lead to revisitng the concept of
reproducible research. This concept is based on the premises that an
independent researcher/analyst is enabled to reproduce the reported
results. Reproducibility covers therefore a spectrum and mix of method,
data, and analytical code/scripts per se openly available for scrutiny
by others.

Independent (operational) ANS performance builds on transparent and
robust performance monitoring and reporting. Within the European context
EUROCONTROL's Performance Review Unit (PRU) is committed to provide
impartial analyses of the performance of the European Air Navigation
System based on verified data. With a view to enhance transparency, PRU
is - inter alia - making performance-related data publicly available via
its Portal, \url{http://ansperformance.eu}, since early 2017 combined
with the methodological documentation for the calculation of the
performance indicators used under the EUROCONTROL Performance Revew
System and the European Union Performance Scheme. There has been
recognition of using common and openly available tools and data in ATM
research and operational validation. With the advent of crowd sourced
data and aviation databases there is an opportunity to advance the state
of the art by combining open data and government organisation collected
data to establish a curated platform supporting the quest for
reproducibility in ANS Performance. One of the associated building
blocks is the performance reference trajectory
{[}\protect\hyperlink{ref-spinielli_2018}{5}{]}. The performance
reference trajectory will be deployed as Open Data throughout 2019. The
trajectory project builds on crowd sourced surveillance data fused with
correlated position reports collected from European air navigation
service providers. Based on the Open Data model it will be possible to
augment the data processing (e.g.~data coverage, time synchronisation),
establishment of derived data bases (e,g fleet data base), associated
performance models (e.g.~aircraft performance model), but also the
continual further development and validation of operational performance
measures/indicators.

This paper builds on the previous work and implements the
reproducibility paradigm. As a use-case example, a fully reproducible
paper is devised assessing the performance in the arrival phase for 2
European airports. The underlying data, supporting data transformation
and analysis, and the paper ``source text'' are made openly available at
the paper companion web repository
\url{https://github.com/euctrl-pru/reproducible-ans-performance-paper}.
This enables the independent review and validation of the work reported.
Conceptually, an interested researcher/analyst can reproduce this paper
by using the same toolset or validate the results by using the source
data or interim data products if the complete reproduction of the study
would be too time/resource consuming.

The contributions of this paper are as follows:

\begin{itemize}
\tightlist
\item
  first fully reproducible ANS Performance related paper: The paper is
  written in RMarkdown allowing for the analysis of how text, code
  chunks, and visualisations are produced (including associated
  parameter settings) on the basis of the ``source text''.
\item
  transparent data analyses and result generation through supporting
  web-page: All data used in this paper, intermediate data artifacts,
  and results including the associated transformation are publicly
  available for inspection, validation, or reproduction.
\item
  use-case analysis of ANS Performance during the arrival phase for 2
  European airports: As an example the classical `additional ASMA time'
  performance indicator is implemented on the basis of the provided data
  with showcasing trajectory derived capabilities to further study the
  efficiency of the arrival phase (e.g.~impact of sequencing strategies
  {[}stack holdings, point merge{]}).
\end{itemize}

\section{Background -
Reproducibility}\label{background---reproducibility}

@@

The terms replication and reproducibility are often used interchangeably
and there is a debate about the necessary and sufficient conditions for
these terms. Results are considered replicable, ``if there is sufficient
information available for independent researchers to make the same
{[}consistent{]} findings using the same procedures with new data''
({[}\protect\hyperlink{ref-gandrud_2015}{6}{]}, p.~3f). Replicability,
thus, does not aim to produce the ``same'' (aka identical) results. Here
the goal is to establish confirmation of the approach and findings by
deriving same conclusions.\footnote{For example, the same data
  collection procedures may not necessarily result in ``identical''
  results as the time horizon might be different, (new or different)
  data for another study case show different result values, but the
  conclusions are consistent with the replicated approach.} Accordingly,
results are reproducible, if an independent researcher can recreate the
findings based on the given information. In our domain (i.e.~operational
ANS Performance and associated data analysis) this will require the
access to data and derived summaries, how associated visualisations have
been generated, the accompanying explanatory text, and conclusions drawn
for the analysis.

Reproducibility has also to be seen in light of today's ``publish or
perish'' context. It appears that higher emphasis is given to the
frequency of publication than the validity of the produced results.
Within the medical world cases like the Duke ovarian cancer scandal
{[}\protect\hyperlink{ref-kolata_2011}{1}{]} are not single incidents,
but also demonstrate how errors during the data analysis - in this case
during the data preparation phase using spread sheets - can discredit
the work. Despite the increasing awareness, the adaption of
reproducibility or moving to other open scholarship/investigation
techniques is not yet achieved. This can be contrasted by emerging
practices like data journalism. One interesting example is buzzFeed's
article on the use of surveillance planes operated by law enforcement
and military for a variety of purposes. The authors were able to
demonstrate based on Open Data that these purposes range from tracking
drug trafficking to testing new surveillance technology
{[}\protect\hyperlink{ref-peteraldhous_2016}{7}{]}. The credibility of
the article comes from the fact that all artifacts are made available to
reproduce the findings or study the chosen approach and data
transformations made leading to the conclusions of the article.

@@

The reproducibility discussion focuses strongly on the recreation of
results by other researchers/analysts. A related problem is the
re-generating of already completed (and potentially published) results
with the - previously - available same input data. This is a fundamental
issue that can be summarised as ``can a future me re-do what I had done
in the past''? This problem space ranges across the complete data
analysis life-cyle, including archiving the orignially used input data
or the steps to generate the study data, documenting the data cleaning
and transformation steps including the parameters chosen, etc. By
definition such a capability is a precondition for reproduction or
replication by others. This implies that the analytical steps including
the steps to produce summaries or estimates, visualisations, and
associated conclusions are documented and archived. In order to enable
others to recreate the own findings, such steps and documentation would
need to be shareable. This consideration has a strong bearing on the
toolset chosen for this paper.

Some of the reasons why researchers/analysts do not publish data and
computational procedures that allow reproduction will be hard to change
{[}\protect\hyperlink{ref-daniel_nust_2018_2361304}{9}{]}. The
predominant technical reasons have been listed above. However, there are
also personal reasons such as the fear for floating ideas or loss of
competitive advantage. Other reasons deal with more technical problems,
e.g.~lack of capabilities, procedures.

Another incentive for the adoption of an open and reproducible research
practice can be seen in the wider recognition through increased
citations, media attention, discussion with the community and
collaborators, and ultimately funding opportunities. The recent years
have seen a strong use of research blogs and community platforms
supporting the publication of smaller steps or intermediate results. The
combination of web presentation and discussions offers a good basis for
an initial dialogue. Platforms like google scholar or researchgate.net
offer additional features for outreach and communication. This equally
applies for paper achives such as arxiv.org for pre-prints. Based on the
credibility crisis discussion a number of journals are now promoting -
or even requiring - the provision of analytical data in support of the
paper publication. There is also an increasing number of requirements
for national or European funded research with respect to data management
and Open Data access to the research.

@@

\begin{itemize}
\tightlist
\item
  archaeology: Marwick et al. (2017) focusses on computational
  reproducibility in archaeological research by providing basic
  principles. The case study of their implementation uses open source
  software (R/Rmarkdown/git approach)
  {[}\protect\hyperlink{ref-marwick_2017}{10}{]}. The paper advocates
  the publishing (analytical) data in support of the paper. This should
  not be an issue, even if there are constraints on the overall study
  data.
\item
  engineering / signal processing: Vandewalle (2009) discuss the
  motivation and approach to reproducibility in signal processing
  (i.e.~what, why, and how). Also here the case is made to open access
  to the code and data for validation purposes
  {[}\protect\hyperlink{ref-vandewalle_2009}{11}{]}.
\item
  operational ANS Performance monitoring: Spinielli et al. (2018)
  highlight the development and use of Open Flight Trajectories for
  reproducible ANS {[}\protect\hyperlink{ref-spinielli_2018}{5}{]}.
\end{itemize}

@@

\caption{Reproducible research, Fig. 2 in
{[}\protect\hyperlink{ref-marwick_2017}{10}{]}.}\label{fig:reproducible-research}
\end{figure}

Across the literature, researchers have identified a ``reproducibility
spectrum'' ranging from the extreme, i.e.~publication only ( advertising
and marketing of the done work), to almost or even fully replicable
research. The work of Marwick et al. (2017)
{[}\protect\hyperlink{ref-marwick_2017}{10}{]} and Peng (2011)
{[}\protect\hyperlink{ref-Peng1226}{12}{]} uses the analogy of a
spectrum that can be mapped against the scoring of Vandewalle (2009)
describing the level of reproducibility on a 0 to 5 scale
{[}\protect\hyperlink{ref-marwick_2017}{10}{]}. In light of
{[}\protect\hyperlink{ref-leek_2015}{13}{]}, this paper conceptualises
reproducibility as the ``ability to recompute data analytic results
given an observed dataset and knowledge of the data analysis pipeline''
and accordingly ``replicability of a study is the chance that an
independent experiment targeting the same scientific question will
produce a consistent result''.

With Fig. \ref{fig:reproducible-research} and the discussion in this
section, this paper implements a fully reproducible workflow. This is
achieved by establishing this paper with open source software which
allows to combine text, code, and visualisations. The associated source
text can also easily be read with a simple text editor/viewer. From that
perspective, the text is fully inspectable. This includes the majority
of visualisations which are generated based on the data. Accordingly,
the parameters and settings chosen can be validated by an independent
researcher/analyst. Last but not least, the data for this paper is made
available in conjunction with this text as Open Data. This also includes
intermediate data artifacts which allows for the reproduction of the
complete work or portions of it. The conceptual approach and its
building blocks will be presented in more detail in the next section.

\section{Method / Approach}\label{method-approach}

\subsection{Conceptual Approach}\label{conceptual-approach}

@@

The source data for this paper comprises the PRU Performance Reference
Trajectory and operator reported data under the airport operator data
flow. This study also uses some aeronautical information for the use
case airports accessible through AIP or community maintained open data
archives (e.g.~http://openaip.net, http://ourairports.com). These data
flows are processed and stored in the EUROCONTROL data Warehouse. The
approach recognises that the full source data may or cannot be provided
for a variety of reasons. These include volume of the source data sets,
technical constraints allowing access to the source data storage, and
data sharing policies or use limitations. For the latter cases, PRU
promotes the sharing of a so-called representative sample. The idea is
that with the provision of the associated processing code, other
researchers/practitioners are able to reconstruct a part of the analytic
data - and subsequently part of the reported results - on the basis of
the provided source data set.

Analytical data refers to the data sets established based on the
aforementioned processing code. An essential defining characteristic of
analytic data is that it is the data basis on which the reported results
can be produced. The abstraction between source data and analytic data
allows for the sharing of such data as limitations are typically only
appliable to the source data. We follow the notion of
{[}\protect\hyperlink{ref-marwick_2017}{10}{]} by referring to the
provision of a representative sample of 8 days accompanying this paper
as source data.

@@

\subsection{Toolset}\label{toolset}

This work builds on the RStudio tools for the R language
{[}\protect\hyperlink{ref-rcoreteam_2018}{14}{]} including Git (and the
web-based repository managers GitHub and GitLab) as underlying version
control system. The R language was originally developed within the
statistical community supporting the task of statistical reporting by
providing routines for the statistical computing and visualisation.
Being open source, the R community is actively engaging and sharing
related software packages to enhance the core functionality. Without
limiting the impact of other packages, the development of knitr
{[}\protect\hyperlink{ref-R-knitr}{15}{]} and RMarkdown
{[}\protect\hyperlink{ref-xie2018}{16}{]}, ggplot
{[}\protect\hyperlink{ref-wickham_2016}{17}{]} for visualisation, and
the so-called tidyverse packages and RStudio IDE
{[}\protect\hyperlink{ref-rstudioteam_2015}{18}{]} provide the basis for
an open source data analysis ecosystem.

>>>>

\hypersetup{
            pdftitle={Towards Reproducibility in ANS Performance},
            pdfauthor={Enrico Spinielli, Rainer Koelle},
            pdfkeywords={reproducibility, ANS performance, flight trajectory, open data, performance indicator},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Towards Reproducibility in ANS Performance}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Air transportation is undergoing a fundamental transformation. On the political, strategic, and operational level work is directed to meet the future growth of air traffic.
To meet this goal, operational concepts and technological enablers are deployed with the promises of distinct performance improvements. These improvements are typically widely marketed, however, the underlying evidence is not made available. From that respect Air Navigation System Performance faces varying levels of transparency as seen in other industries.
This paper implements the reproducibility paradigm by providing data, code, and results openly to enable another researcher or analyst to reproduce - and potentially - improve or build on our work. This work is part of a wider thread to establish Open Data and Open Source based data analytical capabilities for Air Navigation System Performance in Europe.
The implementation of the paradigm is demonstrated on the basis of the performance reference trajectory by replicating a classical performance measure, and then moving into investigating the application of trajectory based information for enhancing the current state of performance monitoring for the arrival phase. This paper applies the concept and approach as a use case analysis for 2 European airports with different strategies for sequencing arrival traffic (i.e.~stack holding and point merge).
The research reported demonstrates the feasibility of the reproducibility approach. This work introduces a novel workflow by making available a companion web repository that provides data, code, and instructions to fully reproduce the paper.
\end{abstract}

% keywords
\begin{IEEEkeywords}
reproducibility; ANS performance; flight trajectory; open data; performance indicator
\end{IEEEkeywords}

@@

% no keywords

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

To meet the future growth of air traffic, political decision makers, strategic planers, practitioners and researchers aim to enhance the performance of air navigation services (ANS).
Typically, deployment programmes are revolving around public-private partnerships and regional funding programmes resulting in a wide mix of consortia and stakeholder subsets.
The actual deployment of novel concepts or technological enablers is then driven by the programmes and to a lesser extent by operational network wide priorities following a piece meal approach.

Moreover, the success of these deployments is typically supported by promising business cases and operational performance improvements.
While these improvements are marketed, the underlying data and associated analyses in support of the results are seldomly made available.
Explanations range from data sensitivity, commercial interests and property rights, through to complexity of the data analyses or simulations, and the associated volume of data.
Similar practise is reported in other domains, e.g.~medicine {[}\protect\hyperlink{ref-kolata_2011}{1}{]}, psychology {[}\protect\hyperlink{ref-ritchie_2012}{2}{]}, ecology {[}\protect\hyperlink{ref-fraser_etal_2018}{3}{]}. Fidler and Wilcox {[}\protect\hyperlink{ref-fidler_wilcox_2018}{4}{]} provides a recent review of the reproducibility discussion.
Throughout the recent years, this lead to revisiting the concept of reproducible research.
This concept is based on the premises that an independent researcher/analyst is enabled to reproduce the reported results.
Reproducibility covers therefore a spectrum and mix of method, data, and analytical code/scripts per se openly available for scrutiny by others.

Independent (operational) ANS performance builds on transparent and robust performance monitoring and reporting.
Within the European context, EUROCONTROL's Performance Review Unit (PRU) is committed to provide impartial analyses of the ANS performance.
With a view to enhance transparency, PRU is - inter alia - making performance-related data publicly available via its portal, \url{http://ansperformance.eu}, since early 2017 combined with the methodological documentation for the calculation of the performance indicators used under the EUROCONTROL Performance Review System and the European Union Performance Scheme.
There has been recognition of using common and openly available tools and data in ATM research and operational validation.
With the advent of crowd sourced data and aviation databases there is an opportunity to advance the state of the art by combining open data and government organisation collected data to establish a curated platform supporting the quest for reproducibility in ANS Performance.
One of the associated building blocks is the performance reference trajectory {[}\protect\hyperlink{ref-spinielli_2018}{5}{]}.
At the time of writing, the performance reference trajectory is in its design stage. The trajectory project builds on crowd sourced surveillance data fused with correlated position reports collected from European air navigation service providers.
Based on the Open Data model it will be possible to augment the data processing (e.g.~data coverage, time synchronisation), establish derived data bases (e,g fleet data base), and derive associated performance models (e.g.~aircraft performance model).
This setup enables the continual further development and validation of operational performance measures/indicators.

This paper builds on the previous work and implements the reproducibility paradigm.
As a use-case example, a fully reproducible paper is devised assessing the performance in the arrival phase for 2 European airports.
The underlying data, supporting data transformation and analysis, and the paper ``source text'' are made openly available at the paper companion web repository \url{https://github.com/euctrl-pru/reproducible-ans-performance-paper}.
This enables the independent review and validation of the work reported.

\hypertarget{background---reproducibility}{%
\section{Background - Reproducibility}\label{background---reproducibility}}

\hypertarget{scope-and-definition}{%
\subsection{Scope and Definition}\label{scope-and-definition}}

@@

The terms replication and reproducibility are often used interchangeably and there is a debate about the necessary and sufficient conditions for these terms.
Results are considered replicable, ``if there is sufficient information available for independent researchers to make the same {[}consistent{]} findings using the same procedures with new data'' ({[}\protect\hyperlink{ref-gandrud_2015}{6}{]}, p.~3f).
Replicability, thus, does not aim to produce the ``same'' (aka identical) results. Here the goal is to establish confirmation of the approach and findings by deriving same conclusions.\footnote{For example, the same data collection procedures may not necessarily result in ``identical'' results as the time horizon might be different, (new or different) data for another study case show different result values, but the conclusions are consistent with the replicated approach.}
Accordingly, results are reproducible, if an independent researcher can recreate the findings based on the given information. In our domain (i.e.~operational ANS Performance and associated data analysis) this will require the access to data and derived summaries, how associated visualisations have been generated, the accompanying explanatory text, and conclusions drawn from the analysis.

Reproducibility has also to be seen in light of today's ``publish or perish'' context. It appears that higher emphasis is given to the frequency of publication than the validity of the produced results. Within the medical world cases like the Duke ovarian cancer scandal {[}\protect\hyperlink{ref-kolata_2011}{1}{]} are not single incidents, but also demonstrate how errors during the data analysis - in this case during the data preparation phase using spread sheets - can discredit the work.
Despite the increasing awareness, the adoption of reproducibility or moving to other open scholarship/investigation techniques is not yet achieved.
This can be contrasted by emerging practices like data journalism. One interesting example is buzzFeed's article on the use of surveillance planes operated by law enforcement and military. The authors were able to demonstrate based on Open Data that these purposes range from tracking drug trafficking to testing new surveillance technology {[}\protect\hyperlink{ref-peteraldhous_2016}{7}{]}. The credibility of the article comes from the fact that all artefacts are made available to reproduce the findings or study the approach and data transformations leading to the conclusions of the article.

@@

@@

\begin{itemize}
\tightlist
\item
  archaeology: Marwick et al.~(2017) focusses on computational reproducibility in archaeological research by providing basic principles. The case study of their implementation uses open source software (R/RMarkdown/git approach) {[}\protect\hyperlink{ref-marwick_2017}{9}{]}. The paper advocates the publishing (analytical) data in support of the paper. This should not be an issue, even if there are constraints on the overall study data.
\item
  engineering / signal processing: Vandewalle {[}\protect\hyperlink{ref-vandewalle_2009}{10}{]} discuss the motivation and approach to reproducibility in signal processing (i.e.~what, why, and how). Also here the case is made to open access to the code and data for validation purposes .
\item
  operational ANS Performance monitoring: Spinielli et al.~{[}\protect\hyperlink{ref-spinielli_2018}{5}{]} highlight the development and use of Open Flight Trajectories for reproducible ANS .
\end{itemize}

@@

\caption{Reproducible research, Fig. 2 in {[}\protect\hyperlink{ref-marwick_2017}{9}{]}.}\label{fig:reproducible-research}
\end{figure}

\hypertarget{reproducibility-spectrum}{%
\subsection{Reproducibility Spectrum}\label{reproducibility-spectrum}}

Across the literature, researchers have identified a ``reproducibility spectrum'' ranging from the extreme, i.e.~publication only (advertising and marketing of the done work), to almost or even fully replicable research.
The work of Marwick et al.~{[}\protect\hyperlink{ref-marwick_2017}{9}{]} and Peng {[}\protect\hyperlink{ref-Peng1226}{11}{]} uses the analogy of a spectrum that can be mapped against the scoring of Vandewalle {[}\protect\hyperlink{ref-vandewalle_2009}{10}{]} describing the level of reproducibility on a 0 to 5 scale {[}\protect\hyperlink{ref-marwick_2017}{9}{]}.
In light of {[}\protect\hyperlink{ref-leek_2015}{12}{]}, this paper conceptualises reproducibility as the ``ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline'' and accordingly ``replicability of a study is the chance that an independent experiment targeting the same scientific question will produce a consistent result''.

With Fig. \ref{fig:reproducible-research} and the discussion in this section, this paper implements a fully reproducible workflow.
This is achieved by establishing this paper with open source software which allows to combine text, code, and visualisations.
The associated source text can also easily be read with a simple text editor/viewer. From that perspective, the text is fully inspectable.
This includes the majority of visualisations which are generated based on the data. Accordingly, the parameters and settings chosen can be validated by an independent researcher/analyst. Last but not least, the data for this paper is made available in conjunction with this text as Open Data. This also includes intermediate data artefacts which allows for the reproduction of the complete work or portions of it.
The conceptual approach and its building blocks will be presented in more detail in the next section.

\hypertarget{method-approach}{%
\section{Method / Approach}\label{method-approach}}

\hypertarget{conceptual-approach}{%
\subsection{Conceptual Approach}\label{conceptual-approach}}

@@

The source data for this paper comprises the PRU Performance Reference Trajectory and operator reported data under the airport operator data flow. This study also uses some aeronautical information for the use case airports accessible through AIP or community maintained open data archives (e.g.~http://openaip.net, http://ourairports.com).
These data flows are processed and stored in the EUROCONTROL data warehouse.
The approach recognises that the full source data may or cannot be provided for a variety of reasons.
These include volume of the source data sets, technical constraints allowing access to the source data storage, and data sharing policies or use limitations.
For the latter cases, PRU promotes the sharing of a so-called representative sample.
The idea is that with the provision of the associated processing code, other researchers/practitioners are able to reconstruct a part of the analytic data - and subsequently part of the reported results - on the basis of the provided source data set.

Analytical data refers to the data sets established based on the aforementioned processing code.
An essential defining characteristic of analytic data is that it is the data basis on which the reported results can be produced.
The abstraction between source data and analytic data allows for the sharing of such data as limitations are typically only applicable to the source data. We follow the notion of {[}\protect\hyperlink{ref-marwick_2017}{9}{]} by referring to the provision of a representative sample of 8 days accompanying this paper as source data.

@@

\hypertarget{toolset}{%
\subsection{Toolset}\label{toolset}}

This work builds on the RStudio tools for the R language {[}\protect\hyperlink{ref-rcoreteam_2018}{13}{]} including Git (and the web-based repository managers GitHub and GitLab) as underlying version control system.
The R language was originally developed within the statistical community supporting the task of statistical reporting by providing routines for the statistical computing and visualisation.
Being open source, the R community is actively engaging and sharing related software packages to enhance the core functionality.
Without limiting the impact of other packages, the development of knitr {[}\protect\hyperlink{ref-R-knitr}{14}{]} and RMarkdown {[}\protect\hyperlink{ref-xie2018}{15}{]}, ggplot {[}\protect\hyperlink{ref-wickham_2016}{16}{]} for visualisation, and the so-called tidyverse packages and RStudio IDE {[}\protect\hyperlink{ref-rstudioteam_2015}{17}{]} represent an open source ecosystem for data analysis.
A key feature for the implementation of the reproducibility paradigm is the fact that RMarkdown supports the combination of text, analytical code, and visualisations in a single document.
A minimal example of an RMarkdown document, a plain-text file with the conventional
extension \texttt{.Rmd} is shown below:
